---
title: "Assignment 1"
author: "Qian He"
date: "8/26/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(plot.matrix)
library(pracma)
library(ggplot2)
library(reshape)
library(resample)
library(MASS)
```

## Question 1

### 1
```{r}
par(mfrow = c(3, 2))

TC_1 <- rep(c(rep(1, 15), rep(0, 15)), 8)[1:240]
TC_1_s <- (TC_1 - mean(TC_1)) / sd(TC_1)
plot(TC_1_s, type = "l")

TC_2 <- rep(c(rep(0, 25), rep(1, 20)), 6)[1:240]
TC_2_s <- (TC_2 - mean(TC_2)) / sd(TC_2)
plot(TC_2_s, type = "l")

TC_3 <- rep(c(rep(1, 25), rep(0, 35)), 4)[1:240]
TC_3_s <- (TC_3 - mean(TC_3)) / sd(TC_3)
plot(TC_3_s, type = "l")

TC_4 <- rep(c(rep(1, 15), rep(0, 25)), 6)[1:240]
TC_4_s <- (TC_4 - mean(TC_4)) / sd(TC_4)
plot(TC_4_s, type = "l")

TC_5 <- rep(c(rep(1, 20), rep(0, 20)), 6)[1:240]
TC_5_s <- (TC_5 - mean(TC_5)) / sd(TC_5)
plot(TC_5_s, type = "l")

TC_6 <- rep(c(rep(1, 25), rep(0, 15)), 6)[1:240]
TC_6_s <- (TC_6 - mean(TC_6)) / sd(TC_6)
plot(TC_6_s, type = "l")
```

### 2
```{r}
TC <- matrix(c(TC_1_s, TC_2_s, TC_3_s, TC_4_s, TC_5_s, TC_6_s), 240, 6)

CM_TC <- cor(TC)

plot(CM_TC)
```

### 3
```{r}
par(mfrow = c(3, 2))

SM_1 <- matrix(rep(0), 21, 21)
SM_1[2:6, 2:6] <- 1
plot(SM_1)

SM_2 <- matrix(rep(0), 21, 21)
SM_2[2:6, 15:19] <- 1
plot(SM_2)

SM_3 <- matrix(rep(0), 21, 21)
SM_3[8:13, 2:6] <- 1
plot(SM_3)

SM_4 <- matrix(rep(0), 21, 21)
SM_4[8:13, 15:19] <- 1
plot(SM_4)

SM_5 <- matrix(rep(0), 21, 21)
SM_5[15:19, 2:6] <- 1
plot(SM_5)

SM_6 <- matrix(rep(0), 21, 21)
SM_6[15:19, 15:19] <- 1
plot(SM_6)

tmpSM <- array(c(SM_1, SM_2, SM_3, SM_4, SM_5, SM_6), c(21, 21, 6))
```

```{r}
SM <- matrix(tmpSM, 6, 441, byrow=T)

CM_SM <- cor(t(SM))

plot(CM_SM)
```

### 4
```{r}
t <- matrix(rnorm(240*6, mean = 0, sd = sqrt(0.25)), 240 ,6)
plot(cor(t))

s <- matrix(rnorm(6*441, mean = 0, sd = sqrt(0.015)), 6 ,441)
plot(cor(t(s)))
```

```{r}
hist(t, probability = T)
curve(dnorm(x, mean = 0, sd = sqrt(1.96*0.25)), add = T)

hist(s, probability = T)
curve(dnorm(x, mean = 0, sd = sqrt(1.96*0.015)), add = T)
```

```{r}
t_s <- t%*%s

plot(cor(t_s), border = NA)
```

### 5
```{r}
X <- (TC + t)%*%(SM + s)

dim(TC%*%s)
dim(t%*%SM)
```

```{r}
ts_100 <- melt(data.frame(n=1:240, X[,sample(1:240, 100, replace = F)]), id.vars = "n")

ggplot(ts_100, aes(x=n, y=value, col=factor(variable))) + geom_line()
```

```{r}
plot(colVars(X))
```

```{r}
X_s <- scale(X)
```

## Question 2

### 1
```{r}
A_LSR <- solve(t(TC)%*%TC, t(TC)%*%X_s)
D_LSR <- X_s%*%t(A_LSR)

par(mfrow = c(1, 2))
plot(matrix(A_LSR[1, ], 21, 21), border = NA)
plot(D_LSR[, 1], type = "l")
```

```{r}
plot(D_LSR[, 3], X_s[, 30])
plot(D_LSR[, 4], X_s[, 30])
```

The reason is that the increment vector is 60 which is different from each other. At the same time, since there is a very clear relationship which means when TC is used to build the 30th column of X_s and the products are the each row of (TC + t) and 30th column of (SM + s) which is 6X6 for each product, the most important point is that the 30th column of SM is t([0, 0, 1, 0, 0, 0]) which means only the third column of TC is used to compute the 30th column of X_s. Since (TC + t) and (SM + s) are used, so there are some noises.

### 2
```{r}
V <- 441
lamda <- 0.8
lamda_hat <- lamda * V

I <- diag(1, 6, 6)

A_RR <- solve(t(TC)%*%TC + lamda_hat*I, t(TC)%*%X_s)
D_RR <- X_s%*%t(A_RR)
```

```{R}
correlate <- function(TC, D){
  cor_vector <- array(0, 6)
  for (i in 1:ncol(TC)) {
    cor_vector[i] <- cor(TC[, i], D[, i])
  }
  return (cor_vector)
}

C_TLSR <- correlate(TC, D_LSR)
C_TLSR

C_TRR <- correlate(TC, D_RR)
C_TRR
```

```{r}
sum(C_TLSR)
sum(C_TRR)
```

```{r}
lamda_2 <- 1000
lamda_2_hat <- lamda_2 * V

A_RR_2 <- solve(t(TC)%*%TC + lamda_2_hat*I, t(TC)%*%X_s)

plot(A_RR_2[1, ], A_LSR[1, ])
```

### 3
```{r}
rho <- seq(0, 1, 0.05)
N <- 240
x1 <- 21
x2 <- 21
nsrcs <- 6
MSE <- array(0, 21)

for (j in 1:length(rho)) {
  step <- 1/(norm(TC %*% t(TC)) * 1.1)
  thr <- rho[j]*N*step
  Ao <- matrix(0, nsrcs, 1)
  A <- matrix(0, nsrcs, 1)
  Alr <- matrix(0, nsrcs, x1*x2)
  mse <- array(0, 10)
  
  for (a in 1:10) {
    t_new <- matrix(rnorm(240*6, mean = 0, sd = sqrt(0.25)), 240 ,6)
    s_new <- matrix(rnorm(6*441, mean = 0, sd = sqrt(0.015)), 6 ,441)
    
    X_new <- scale((TC + t_new)%*%(SM + s_new))
    
    for (k in 1:(x1*x2)) {
      A <- Ao+step*(t(TC) %*% (X_new[,k]-(TC%*%Ao)))
      A <- (1/(1+thr)) * (sign(A)*pmax(replicate(nsrcs, 0), abs(A)-thr))
    
      for (i in 1:10) {
        Ao <- A
        A <- Ao+step * (t(TC)%*%(X_new[,k]-(TC%*%Ao)))
        A <- (1/(1+thr)) * (sign(A)*pmax(replicate(nsrcs, 0), abs(A)-thr))
      }
      Alr[,k] <- A
    }
    
    Dlr <- X_new%*%t(Alr)
    mse[a] <- sum(sum((X_new-Dlr%*%Alr)^2))/(N*V)
  }
  
  MSE[j] <- mean(mse)
}
```

```{r}
rho
MSE
plot(rho, MSE)
```

### 4
```{r}
rho_best <- 0.6
step <- 1/(norm(TC %*% t(TC)) * 1.1)
thr <- rho_best*N*step
Ao <- matrix(0, nsrcs, 1)
A <- matrix(0, nsrcs, 1)
Alr <- matrix(0, nsrcs, x1*x2)

for (k in 1:(x1*x2)) {
  A <- Ao+step*(t(TC) %*% (X_s[,k]-(TC%*%Ao)))
  A <- (1/(1+thr)) * (sign(A)*pmax(replicate(nsrcs, 0), abs(A)-thr))

  for (i in 1:10) {
    Ao <- A
    A <- Ao+step * (t(TC)%*%(X_s[,k]-(TC%*%Ao)))
    A <- (1/(1+thr)) * (sign(A)*pmax(replicate(nsrcs, 0), abs(A)-thr))
  }
  Alr[,k] <- A
}

Dlr <- X_s%*%t(Alr)
```

```{r}
C_TRR <- correlate(TC, D_RR)
C_TRR

C_SRR <- correlate(t(SM), t(A_RR))
C_SRR

C_TLR <- correlate(TC, Dlr)
C_TLR

C_SLR <- correlate(t(SM), t(Alr))
C_SLR
```

```{r}
sum(C_TRR)
sum(C_SRR)
sum(C_TLR)
sum(C_SLR)
```

```{r}
plot(cor(t(A_RR), t(Alr)))

plot(cor(D_RR, Dlr))
```

### 5
```{r}
Z <- svd(TC, 6)$u
E_V <- svd(TC, 6)$v

plot(E_V)
```

```{r}
par(mfrow = c(2, 2))

plot(Z[, 1], TC_1)
plot(TC_1, type = "l")

plot(Z[, 2], TC_2)
plot(TC_2, type = "l")
```

```{r}
par(mfrow = c(2, 2))

plot(Z[, 3], TC_3)
plot(TC_3, type = "l")

plot(Z[, 4], TC_4)
plot(TC_4, type = "l")
```

```{r}
par(mfrow = c(2, 2))

plot(Z[, 5], TC_5)
plot(TC_5, type = "l")

plot(Z[, 6], TC_6)
plot(TC_6, type = "l")
```

```{r}
rho_PCR <- 0.001
step <- 1/(norm(TC %*% t(TC)) * 1.1)
thr <- rho_PCR*N*step
Ao <- matrix(0, nsrcs, 1)
A <- matrix(0, nsrcs, 1)
A_PCR <- matrix(0, nsrcs, x1*x2)

for (k in 1:(x1*x2)) {
  A <- Ao+step*(t(TC) %*% (X_s[,k]-(TC%*%Ao)))
  A <- (1/(1+thr)) * (sign(A)*pmax(replicate(nsrcs, 0), abs(A)-thr))

  for (i in 1:10) {
    Ao <- A
    A <- Ao+step * (t(TC)%*%(X_s[,k]-(TC%*%Ao)))
    A <- (1/(1+thr)) * (sign(A)*pmax(replicate(nsrcs, 0), abs(A)-thr))
  }
  A_PCR[,k] <- A
}

D_PCR <- X_s%*%t(A_PCR)
```

```{r}
par(mfrow = c(3, 2))

for (num in 1:6) {
  plot(D_PCR[num, ], A_PCR[, num])
}
```

```{r}
par(mfrow = c(3, 2))

for (num in 1:6) {
  plot(D_LSR[num, ], A_LSR[, num])
}
```

```{r}
par(mfrow = c(3, 2))

for (num in 1:6) {
  plot(D_RR[num, ], A_RR[, num])
}
```

```{r}
par(mfrow = c(3, 2))

for (num in 1:6) {
  plot(Dlr[num, ], Alr[, num])
}
```